A multi-dimensional percolation approach to characterize sustainable mega-city regions

### 2018-10-17 Docs/Marami2018/Communication/marami2018.tex
slides

@@

@@

\textbf{Research objective : } \textit{}

@@

\sframe{Formalization}{

@@

@@

%We apply the heuristic to urban morphology and road network topology measures in Europe. More precisely, a grid with resolution 50km of population density morphology indicators and road network topology indicators, has been computed on spatial moving windows for all European Union by \cite{raimbault2018urban}. We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.

@@

\sframe{Experience plan}{

@@

% + explain openmole

@@

@@

@@

\sframe{Results: Pareto fronts}{

@@

@@

\sframe{Extrapolating transportation flows}{

@@

\sframe{Calibration}{

@@

@@

$\rightarrow$ 
\smallskip

@@

@@

>>>>

$\rightarrow$ Need to combine morphological and functional dimensions of cities \cite{burger2012form}

@@

$\rightarrow$ Interaction between networks and territories to capture the form-function link \cite{raimbault2018caracterisation}; potential application to sustainability of urban systems

@@

\textbf{Research objective : } \textit{Investigate a multi-dimensional percolation of territorial networks taking into account urban morphology and road network topology; endogenous characterization of urban regions.}

@@

\sframe{Multilayer percolation}{

@@

\textit{Multi-dimensional network percolation heuristic, similar to multilayer percolation} \cite{boccaletti2014structure}

\centering
\includegraphics[width=0.8\textwidth]{figures/principle.pdf}

\raggedright

\textbf{Parameters: } percolation radius $r_0$, percolation thresholds $\theta_i$ for each layer

@@

%We apply the heuristic to urban morphology and road network topology measures in Europe. More precisely, a grid with resolution 50km of population density morphology indicators and road network topology indicators, has been computed on spatial moving windows for all European Union by \cite{raimbault2018urban}.

\textit{Territorial indicators computed for Europe by \cite{raimbault2018urban}}

% here add maps to better visualize the spatial fields

\includegraphics[width=0.49\textwidth]{figures/indics_morpho_discrquantiles.png}
\includegraphics[width=0.49\textwidth]{figures/indics_network_en_areasize100_offset50_factor0_5.png}

\textbf{Population distribution morphology} and \textbf{Network topology} (betweenness, closeness, clustering, efficiency, \ldots) computed on 50km spatial windows (Eurostat density grid and OpenStreetMap)

@@

\sframe{Network construction}{

%We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.

\textbf{Two layers: } population density (threshold $\theta_P$) and network characteristics (threshold $\theta_N$) taken among \{Number of edges, Number of vertices, Cyclomatic number $\mu$, Euclidian efficiency $v$ \}; percolated with a radius $r_0$

% banos2012towards

\textbf{Rationale: } \textit{two locations will be in relation if they are close, have a high population density and given network characteristics.}

% note : euclperf should be minimized whereas other maximized : issue ?

\textbf{Implementation: } construction of a single layer spatial network given the condition on the two layers and distances; extraction of connected components.

\textbf{Experience plan: } grid sampling for $r_0,\theta_P,\theta_N$ and network variables; additional gravity potential parameters $\gamma,d_0$ (detailed after)

$\rightarrow$ 4800 parameter points

@@

% + explain openmole -> later in potential developments

@@

\vspace{-0.4cm}
\textit{Extraction of endogenous polycentric mega-city regions \cite{hall2006polycentric}}

% note : forgot r_0 in figs
\includegraphics[width=\textwidth,height=0.9\textheight]{figures/totalPop4183694_00056402_ecount850_radius8000.png}

\sframe{Different endogenous morphologies}{

% TODO for the paper : check effect of adding network threshold : methodo info on multilayer percol in itself

\hspace{-0.5cm}
\includegraphics[width=0.52\textwidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}
\includegraphics[width=0.52\textwidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}

@@

% explain indicators : convex hull.

\textbf{Application: } sustainability indicators for the endogenous urban regions; proxys for two conflicting dimensions: GHG emissions and economic integration \cite{viguie2012trade}.

\textbf{Data: } EDGAR database for GHG emissions (v4.3.2)

\cite{janssens2017edgar}

\textbf{Estimation: } Abstract flows approximated with a gravity model

\[
\phi_{ij} = \left(\frac{v_i v_j}{(\sum_k v_k)^2}\right)^\gamma \cdot \exp\left(\frac{-d_{ij}}{d_0}\right)
\]

where $v_k$ are either effective local GHG emissions or population (economic activity scaling law of population \cite{bettencourt2007growth})

$\rightarrow$ sum of flows within the geographical span of the cluster (convex hull) approximate potential emissions and economic activity

@@

\sframe{Pareto fronts for sustainability}{

@@

\textit{Superposing Pareto front for observed population and emissions, on all clusters.}

\includegraphics[height=0.8\textheight]{figures/full_effective_pareto.png}

% -> full fronts in reserve slide

@@

\sframe{Pareto fronts for aggregated indicators}{

\textit{Aggregated sustainability indicators suggest some configurations are more Pareto efficient (high $\gamma$ regime, activities with high added value).}

\includegraphics[height=0.8\textheight]{figures/aggreg_paretos_radiuspopthq.png}

@@

\sframe{An optimal morphology ?}{

%# Cumulative Proportion  0.7296 0.9650 0.99761 1.00000
%                PC1       PC2         PC3         PC4
%moran   -0.3088585 0.9493848 -0.04444327  0.03605266
%avgdist  0.5417362 0.1415668 -0.82239570 -0.10072759
%entropy  0.5108424 0.2140447  0.45847647 -0.69499942
%slope    0.5917502 0.1811415  0.33390034  0.71100630

\textit{Optimal degree of monocentricity of the system for emissions.}

\includegraphics[height=0.85\textheight]{figures/aggreg_morpho_pc1-emissions.png}

\sframe{Morphological trade-offs}{

\vspace{-0.3cm}

\textit{No ``optimal'' cities, different forms yield different compromises in terms of relative indicators.}

\includegraphics[height=0.85\textheight]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale.png}

\sframe{Towards a systematic calibration}{

@@

Grid sampling to explore regions rapidly limited

$\rightarrow$ towards the use of genetic algorithms on grid, made smooth with the OpenMOLE software \url{https://next.openmole.org/}

\includegraphics[height=0.35\textheight]{figures/openmole.png}

\raggedright\justify

\footnotesize \textit{OpenMOLE: (i) embed any model as a black box; (ii) transparent access to main High Performance Computing environments; (iii) model exploration and calibration methods.}

\textbf{Apply to the summer school !} \texttt{https://exmodelo.org/}

@@

$\rightarrow$ Transfer to policy making recommandations: Pareto-optimal configuration can be used for the planning of regional transportation networks, subsidies policies, etc.
\smallskip

@@

$\rightarrow$ Extrapolation of transportation flows

@@

\sframe{Results: effective emissions}{

\textit{Effective emissions exhibit a supralinear scaling of population}

\includegraphics[height=0.9\textheight]{figures/aggreg_effective.png}

\sframe{Results: all clusters Pareto fronts}{

\textit{Variation of Pareto front patterns when potential parameter $\gamma,d_0$ vary.}

\includegraphics[width=0.9\textwidth]{figures/full_paretos.png}

\sframe{Results: an optimal morphology}{

\textit{More monocentric areas are more optimal in terms of relative emissions and efficiency ?}

\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relefficiency.png}
\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relemissions.png}

% TODO : curves for emissions rescale to the same -> link with Caruso profiles ? ; investigate that, and why not for efficiency.
